Efficiently stable presentations from error-correcting codes

### 2023-10-05 efficient_stability.tex

\author[1]{Michael Chapman}
\author[2]{Thomas Vidick}
\author[3]{Henry Yuen}
\affil[1]{}
\affil[2]{ Faculty of Mathematics and Computer Science, Weizmann Institute of Science, Israel}
\affil[3]{}

@@

\begin{definition}[Stability]\label{def:eff-stab-intro}
Let $G = \langle S:R\rangle $ be a finitely presented group, $\mu_S$ a distribution on $S$ and $\mu_R$ a distribution on $R$. For $\delta:[0,1]\to[0,1]$ such that $\lim_{t\to 0}\delta(t)=0$ and an integer $d\geq 1$ we say that the presentation $G=\langle S:R\rangle$ is $(\delta,\mu_S,\mu_R,d)$-stable if for every $(\eps,\mu_R)$-almost homomorphism $\phi: \mF(S) \to \mU(\C^d)$ there is a unitary representation $\psi: G \to \mU(\C^d)$ \hnote{there's some mismatch, is it $\mU(\C^d)$ or $\mU_d(\C)$?}\tnote{Let's use the former, I made the correction here} such that
\[ \Es{s\sim \mu_S} \big\|  \phi(s) - \psi(s) \big \|_f^2 \,\leq\, \delta(\eps)\;.\]
We refer to any function $\delta$ satisfying the above as a ``modulus of stability'' of the presentation.
\end{definition}

%\begin{remark}
%\hnote{promoted this from footnote to remark}\tnote{Can we cancel this change? It's a remark below, see Remark~\ref{rk-mus}. But in the intro, it doesn't seem so crucial to highlight. \hnote{Agreed that it's not so crucial; perhaps nix it from the intro?}} It is natural, and will always be the case for us, to take $\mu_S$ as the ``marginal'' of $\mu_R$, i.e.\ the distribution on $S$ obtained by sampling $r\sim\mu_R$ and returning a uniformly random $s$ that appears in $r$, when $r$ is considered as a multi-set.
%\end{remark}

@@

\begin{theorem*}[Main, informal]
For every integer $k\geq 1$ there is a presentation $\Z_2^k = \langle S_k:R_k\rangle$ such that $|S_k|,|R_k| = 2^{\poly\log(k)}$ and furthermore this presentation is $(\delta,\mu_{S_k},\mu_{R_k},d)$-stable for all $d \in \N$, where $\delta(\eps)=\poly(\log k,\eps)$, $\mu_{S_k}$ is uniform over the generators and $\mu_{R_k}$ is some distribution over the relations. %\hnote{Spelled the conditions out because if we just let $\mu_S,\mu_R$ be anything, then we get a trivial statement if $\mu_S$, say, is just supported on one generator and $\mu_R$ is just on one relation.}
\end{theorem*}

>>>>

\author[1]{Michael Chapman}
\author[2]{Thomas Vidick}
\author[3]{Henry Yuen}
\affil[1]{}
\affil[2]{ Faculty of Mathematics and Computer Science, Weizmann Institute of Science, Israel}
\affil[3]{Department of Computer Science, Columbia University, USA}

@@

\begin{definition}[Stability]\label{def:eff-stab-intro}
Let $G = \langle S:R\rangle $ be a finitely presented group, $\mu_S$ a distribution on $S$ and $\mu_R$ a distribution on $R$. For $\delta:[0,1]\to[0,1]$ such that $\lim_{t\to 0}\delta(t)=0$ and an integer $d\geq 1$ we say that the presentation $G=\langle S:R\rangle$ is $(\delta,\mu_S,\mu_R,d)$-stable if for every $(\eps,\mu_R)$-almost homomorphism $\phi: \mF(S) \to \mU(\C^d)$ there is a unitary representation $\psi: G \to \mU(\C^d)$ such that
\[ \Es{s\sim \mu_S} \big\|  \phi(s) - \psi(s) \big \|_f^2 \,\leq\, \delta(\eps)\;.\]
We refer to any function $\delta$ satisfying the above as a ``modulus of stability'' of the presentation.
\end{definition}

@@

\begin{theorem*}[Main, informal]
For every integer $k\geq 1$ there is a presentation $\Z_2^k = \langle S_k:R_k\rangle$ such that $|S_k|,|R_k| = 2^{\poly\log(k)}$ and furthermore this presentation is $(\delta,\mu_{S_k},\mu_{R_k},d)$-stable for all $d \in \N$, where $\delta(\eps)=\poly(\log k,\eps)$, $\mu_{S_k}$ is uniform over the generators and $\mu_{R_k}$ is some distribution over the relations. 
\end{theorem*}
